Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-26 drafts/hybrid_zone.tex
heatmaps

%\usepackage[hidelinks]{hyperref}

@@

\paragraph{PDE for linked clines}
We can use this diffusion model for lineages to compute expected clines in ancestry,
i.e., the expected proportion of individuals who inherit from species $A$,
as a function of space, time, and position on the genome.
Precisely, we need the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$,
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
We denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
(Recall that $Z$ depends implicitly on $r$.)
The description above implies that $q_z$ solves the following Kolmogorov backward equation:
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            r p(x,t) (q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} 
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
applied to a backwards-time process!}

@@

%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
%   which we have done using both theory and simulation.
%   Our most useful observations are that clines establish over time $1/s$,
%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
%    tend to move back towards the side where they are common
%   at speed $\sigma/\sqrt{s}$.
%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
%}

\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
This occurs at speed $\sigma/\sqrt{s}$. Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline. 
This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}

\alisa{\paragraph{Targets of selection}}
Most attempts to identify loci under selection in hybrid zones identify alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
Current genomic approaches can provide an additional rich source of information when applied to the study of hybrid zones. In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture \citep[e.g. ][]{Hellenthal2014,Schumer2014}. In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone. Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.

>>>>

@@

\begin{figure}
    \includegraphics{figs/linked-frequencies}
    \caption{
        \textbf{Probabilities of $A$ ancestry,}
        across space (vertical axis, in units of $\sigma$) 
        and time (horizontal axis, in generations).
        In each plot, color corresponds to the expected frequency of $A$ ancestry
        at a particular location in time and space.
        The selection coefficient is $s=.02$.
        \textbf{Top left:} at the selected site, showing establishment and stabilization of the cline
        on a time scale of $1/s=50$ generations.
        \textbf{Bottom left:} at an unliked site,
        with cline flattening continuing with $\sqrt{t}$.
        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
        on the ancestry at the selected site,
        at different distances from the selected site ($r=.01$, .04, and 0.5 Morgans),
        as described in the text (see definition of $q_z(x,t,r)$).
        See figure \ref{sfig:linked_cline_heatmaps_longtime}
        for the same figure over a longer period of time.
    }
    \label{fig:linked_cline_heatmaps}
\end{figure}

\paragraph{Linked clines}
We can use this diffusion model for lineages to compute expected clines in ancestry,
i.e., the expected proportion of individuals who inherit from species $A$,
as a function of space, time, and position on the genome.
Precisely, we need the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$,
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
We denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
(Recall that $Z$ depends implicitly on $r$.)
The description above implies that $q_z$ solves the following Kolmogorov backward equation:
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            r p(x,t) (q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} 
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
applied to a backwards-time process!}

@@

\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
    which we have done using both theory and simulation.
    Our most useful observations are that clines establish over time $1/s$,
    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
    tend to move back towards the side where they are common
    at speed $\sigma/\sqrt{s}$.
    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
}

Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.  \alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g. ][]{Hellenthal2014}.  \alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention. Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}

### 2016-02-26 drafts/supplement.tex
peter edits
.
heatmaps

\documentclass[12pt,titlepage]{article}

%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
\usepackage{graphicx}
%\usepackage{figcaps}
\usepackage[nomarkers,notablist,nofiglist,tablesfirst]{endfloat}
\usepackage{amsfonts}
\usepackage{sectsty}
%\usepackage{subfigure}
%\usepackage{subcaption}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
%\allsectionsfont{\sffamily}
%% THE MANUSCRIPT TITLE

\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}

\usepackage{bibentry}

\usepackage{xr}
\externaldocument{hybrid_zone}

\newcommand{\alisa}[1]{{\em \color{red} #1}}
\newcommand{\plr}[1]{{\em \color{blue} #1}}
\newcommand{\yb}[1]{{\em \color{magenta} #1}}

\newcommand{\given}{\,\vert\,}
\newcommand{\st}{\,\colon\,}
\renewcommand{\and}{\,\&\,}

\begin{document}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}

@@

\end{document}

>>>>

% \documentclass[12pt,titlepage]{article}
% 
% %% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
% %% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
% \usepackage{graphicx}
% %\usepackage{figcaps}
% \usepackage[nomarkers,notablist,nofiglist,tablesfirst]{endfloat}
% \usepackage{amsfonts}
% \usepackage{sectsty}
% %\usepackage{subfigure}
% %\usepackage{subcaption}
% \usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
% \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% %\allsectionsfont{\sffamily}
% %% THE MANUSCRIPT TITLE
% 
% \usepackage{caption}
% \usepackage[labelformat=simple]{subcaption}
% 
% \usepackage{bibentry}
% 
% \usepackage{xr}
% \externaldocument{hybrid_zone}
% 
% \newcommand{\alisa}[1]{{\em \color{red} #1}}
% \newcommand{\plr}[1]{{\em \color{blue} #1}}
% \newcommand{\yb}[1]{{\em \color{magenta} #1}}
% 
% 
% \newcommand{\given}{\,\vert\,}
% \newcommand{\st}{\,\colon\,}
% \renewcommand{\and}{\,\&\,}
% 
% 
% \begin{document}
% \setcounter{table}{0}
% \renewcommand{\thetable}{S\arabic{table}}
% \setcounter{figure}{0}
% \renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}
    \includegraphics{figs/linked-frequencies-longtime}
    \caption{
        \textbf{Probabilities of $A$ ancestry,}
        across space (vertical axis, in units of $\sigma$) 
        and time (horizontal axis, in generations).
        In each plot, color corresponds to the expected frequency of $A$ ancestry
        at a particular location in time and space.
        The selection coefficient is $s=.02$.
        \textbf{Top left:} at the selected site, showing establishment and stabilization of the cline
        on a time scale of $1/s=50$ generations.
        \textbf{Bottom left:} at an unliked site,
        with cline flattening continuing with $\sqrt{t}$.
        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
        on the ancestry at the selected site,
        at different distances from the selected site ($r=.01$, .04, and 0.5 Morgans),
        as described in the text (see definition of $q_z(x,t,r)$).
        See figure \ref{sfig:linked_cline_heatmaps}
        for the same figure over a shorter period of time.
    }
    \label{sfig:linked_cline_heatmaps_longtime}
\end{figure}

@@

% \end{document}
